### 2016-12-30 paper/paper.tex
Major work done. Fixed bug in multi-tag scenario where the error only used a single tag. Committing some previous work on the paper and some miscellaneous code maintanance.

\documentclass[11pt]{article}
%Gummi|065|=)
\title{\textbf{AprilTrack: A Particle-Filter based AprilTag Tracker}}
\author{Daniel Pfrommer}
\date{}

\begin{document}

\maketitle

\begin{abstract}

	The AprilTrack algorithm proposed in this paper seeks to bridge two existing fields of research in computer vision literature: fiducial detection and monocular object tracking. With a rise in the popularity of fiducials, or artificial markers, as a source of ground-truth position data for mobile robots, there has been a considerable interest in designing both robust and blur-resistant fiducials. This paper attempts to improve on existing fiducials such as the AprilTag marker by augmenting fiducial detection algorithms with a particle-filter based monocular object tracker which reliably tracks a fiducial's position and orientation in six degrees of freedom. This novel particle-filter based algorithm is evaluated on several datasets involving both ego-motion and tag-based-motion and is compared to standalone tag detection algorithms.

\end{abstract}

\section{Introduction}

	The majority of previous literature in fiducial detection and monocular object tracking has been focused single-frame fiducial detection 3

\section{Feedback}

\end{document}